---
title: "Column inventories"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup_obs.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r define_paths, include = FALSE}

# only path_observations needs to be changed to model
path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")


path_preprocessing_model    <-
  paste(path_root, "/model/preprocessing/", sep = "")

```

```{r load_libraries_specific, include = FALSE}
library(marelac)
```

# GLODAPv2.2020

# Data preparation

## Load data

```{r read_files}

# Version_IDs <- list.files(
#   path = "/nfs/kryo/work/jenmueller/emlr_cant/observations",
#   pattern = "v_")[1:34]

Version_IDs_1 <- list.files(
  path = "/nfs/kryo/work/jenmueller/emlr_cant/observations",
  pattern = "v_10")[1:4]

Version_IDs_2 <- list.files(
  path = "/nfs/kryo/work/jenmueller/emlr_cant/observations",
  pattern = "v_20")[1:4]

Version_IDs <- c(Version_IDs_1, Version_IDs_2)
rm(Version_IDs_1, Version_IDs_2)

# Version_IDs <- Version_IDs[1:length(Version_IDs)-1]

for (i_Version_IDs in Version_IDs) {
  # i_Version_IDs <- Version_IDs[1]
  
  print(i_Version_IDs)
  
  path_version_data     <-
  paste(path_observations,
        i_Version_IDs,
        "/data/",
        sep = "")
  
  dcant_inv <-
    read_csv(paste(path_version_data,
                   "dcant_inv.csv",
                   sep = ""))
  
  dcant_inv_mod_truth <-
    read_csv(paste(path_version_data,
                   "dcant_inv_mod_truth.csv",
                   sep = "")) %>% 
    filter(method == "total") %>% 
    select(-method)
  
  dcant_inv_bias <-
    read_csv(paste(path_version_data,
                   "dcant_inv_bias.csv",
                   sep = ""))
  
  dcant_inv <- bind_rows(dcant_inv,
                         dcant_inv_mod_truth)
  
  dcant_inv <- dcant_inv %>% 
    mutate(Version_ID = i_Version_IDs)
  
  dcant_inv_bias <- dcant_inv_bias %>% 
    mutate(Version_ID = i_Version_IDs)
  
  params_local <-
    read_rds(paste(path_version_data,
                   "params_local.rds",
                   sep = ""))
  
  params_local <- bind_cols(
    Version_ID = i_Version_IDs,
    MLR_basins = params_local$MLR_basins,
    tref1 = params_local$tref1,
    tref2 = params_local$tref2,
    gap_filling = params_local$gap_filling,
    rarefication = params_local$rarefication,
    vif_max = params_local$vif_max
  )
  
  # dcant_inv <- full_join(dcant_inv, params_local)
  
  tref <- read_csv(paste(path_version_data,
                         "tref.csv",
                         sep = ""))
  
  params_local <- params_local %>% 
    mutate(median_year_1 = sort(tref$median_year)[1],
           median_year_2 = sort(tref$median_year)[2],
           duration = median_year_2 - median_year_1)
  
  # dcant_inv <- dcant_inv %>% 
  #   mutate(duration = duration)
  
  if (exists("dcant_inv_all")) {
    dcant_inv_all <- bind_rows(dcant_inv_all, dcant_inv)
  }
  
  if (!exists("dcant_inv_all")) {
    dcant_inv_all <- dcant_inv
  }

  if (exists("dcant_inv_bias_all")) {
    dcant_inv_bias_all <- bind_rows(dcant_inv_bias_all, dcant_inv_bias)
  }
  
  if (!exists("dcant_inv_bias_all")) {
    dcant_inv_bias_all <- dcant_inv_bias
  }

  if (exists("params_local_all")) {
    params_local_all <- bind_rows(params_local_all, params_local)
  }
  
  if (!exists("params_local_all")) {
    params_local_all <- params_local
  }
  
  
}

rm(dcant_inv, dcant_inv_bias, dcant_inv_mod_truth,
   params_local, tref)

dcant_inv_all <- full_join(dcant_inv_all,
                           params_local_all)

dcant_inv_bias_all <- full_join(dcant_inv_bias_all,
                                params_local_all)

```

## Format column inventories

```{r filter_standard_inventory_depth}

dcant_inv_all <- dcant_inv_all %>%
  filter(inv_depth == params_global$inventory_depth_standard)

```

# MLR regions

## Column inventories

```{r column_inventory_maps_regions_eras, fig.asp=1}

dcant_inv_all %>%
  filter(data_source %in% c("mod", "obs")) %>%
  group_by(tref1) %>%
  group_split() %>%
  head(1) %>% 
  map(
    ~ p_map_cant_inv(df = .x,
                     var = "dcant",
                     subtitle_text = paste("data_source: observations | period:",
                                           unique(.x$tref1))) +
      facet_grid(MLR_basins ~ data_source)
  )

dcant_inv_ensemble <- dcant_inv_all %>% 
  filter(data_source %in% c("mod", "obs")) %>% 
  group_by(lat, lon, data_source, tref1) %>% 
  summarise(dcant_mean = mean(dcant),
            dcant_sd = sd(dcant),
            dcant_range = max(dcant)- min(dcant)) %>% 
  ungroup()

p_map_cant_inv(df = dcant_inv_ensemble,
                     var = "dcant_mean",
                     subtitle_text = paste("data_source: observations")) +
      facet_grid(tref1 ~ data_source)

p_map_cant_inv(
  df = dcant_inv_ensemble,
  var = "dcant_sd",
  breaks = c(seq(0,4,0.4), Inf),
  subtitle_text = paste("data_source: observations")
) +
  facet_grid(tref1 ~ data_source)

p_map_cant_inv(
  df = dcant_inv_ensemble,
  var = "dcant_range",
  breaks = c(seq(0,8,0.8), Inf),
  subtitle_text = paste("data_source: observations")
) +
  facet_grid(tref1 ~ data_source)

dcant_inv_bias_all %>%
  p_map_cant_inv(var = "dcant_bias",
                 col = "bias",
                 subtitle_text = "data_source: mod - mod_truth") +
  facet_grid(MLR_basins ~ .)

```

## Budgets

### Regional

```{r budgets_regional_regions_eras, fig.asp=1}

dcant_inv_budget_sub <- dcant_inv_budget %>%
  filter(Version_ID %in% Version_IDs[1:8])

dcant_inv_budget_sub %>%
  ggplot(aes(data_source, dcant_total/duration, fill = basin_AIP)) +
  scale_fill_brewer(palette = "PuBuGn") +
  geom_col(col = "black") +
  labs(y = expression(Uptake~rate ~ (PgC ~ yr ^ {-1}))) +
  facet_grid(MLR_basins ~ era_duplication)

```

### Global

```{r budget_global_regions_eras}

dcant_inv_budget_sub %>%
  filter(data_source == "obs") %>% 
  group_by(MLR_basins, era_duplication) %>% 
  summarise(rate = sum(dcant_total/duration)) %>% 
  ungroup() %>% 
  ggplot(aes(era_duplication, rate, fill=MLR_basins)) +
  geom_bar(stat="identity", position=position_dodge(), col="black") +
  scale_y_continuous() +
  labs(y = expression(Uptake~rate ~ (PgC ~ yr ^ {-1})),
       x = "") +
  scale_fill_brewer(palette = "PuBuGn")

rm(dcant_inv_budget_sub)

```


# CANYON-B threshold

## Column inventories

```{r column_inventory_maps_CANYON_B_max}

dcant_inv_all_bias_sub <- dcant_inv_all_bias %>%
  filter(Version_ID %in% Version_IDs[c(1,11,12)])

dcant_inv_all_bias_sub %>% 
  p_map_dcant_inv(var = "obs",
                 col = "divergent",
                 subtitle_text = "data_source: observations") +
  facet_grid(CANYON_B_max ~ .)


dcant_inv_all_bias_sub %>%
  p_map_dcant_inv(var = "dcant_bias",
                 col = "bias",
                 subtitle_text = "data_source: mod - mod_truth") +
  facet_grid(CANYON_B_max ~ .)

rm(dcant_inv_all_bias_sub)

```

## Budgets

### Regional

```{r budgets_regional_CANYON_B_max, fig.asp=1}

dcant_inv_budget_sub <- dcant_inv_budget %>%
  filter(Version_ID %in% Version_IDs[c(1,11,12)]) %>% 
  mutate(CANYON_B_max = as.factor(CANYON_B_max))

dcant_inv_budget_sub %>%
  ggplot(aes(data_source, dcant_total/duration, fill = basin_AIP)) +
  scale_fill_brewer(palette = "PuBuGn") +
  geom_col(col = "black") +
  labs(y = expression(Uptake~rate ~ (PgC ~ yr ^ {-1}))) +
  facet_grid(CANYON_B_max ~ .)

```

### Global

```{r budget_global_CANYON_B_max}

dcant_inv_budget_sub %>%
  filter(data_source == "obs") %>% 
  group_by(CANYON_B_max) %>% 
  summarise(rate = sum(dcant_total/duration)) %>% 
  ungroup() %>% 
  ggplot(aes(NA, rate, fill = CANYON_B_max)) +
  geom_bar(stat="identity", position=position_dodge(), col="black") +
  scale_y_continuous() +
  labs(y = expression(Uptake~rate ~ (PgC ~ yr ^ {-1}))) +
  scale_fill_brewer(palette = "PuBuGn") +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank())

rm(dcant_inv_budget_sub)

```


# C* RMSE threshold

## Column inventories

```{r column_inventory_maps_c_star_rmse_max}

dcant_inv_all_bias_sub <- dcant_inv_all_bias %>%
  filter(Version_ID %in% Version_IDs[c(1,13,14)])

dcant_inv_all_bias_sub %>% 
  p_map_dcant_inv(var = "obs",
                 col = "divergent",
                 subtitle_text = "data_source: observations") +
  facet_grid(c_star_rmse_max ~ .)


dcant_inv_all_bias_sub %>%
  p_map_dcant_inv(var = "dcant_bias",
                 col = "bias",
                 subtitle_text = "data_source: mod - mod_truth") +
  facet_grid(c_star_rmse_max ~ .)

rm(dcant_inv_all_bias_sub)

```

## Budgets

### Regional

```{r budgets_regional_c_star_rmse_max, fig.asp=1}

dcant_inv_budget_sub <- dcant_inv_budget %>%
  filter(Version_ID %in% Version_IDs[c(1,13,14)]) %>% 
  mutate(c_star_rmse_max = as.factor(c_star_rmse_max))

dcant_inv_budget_sub %>%
  ggplot(aes(data_source, dcant_total/duration, fill = basin_AIP)) +
  scale_fill_brewer(palette = "PuBuGn") +
  geom_col(col = "black") +
  labs(y = expression(Uptake~rate ~ (PgC ~ yr ^ {-1}))) +
  facet_grid(c_star_rmse_max ~ .)

```

### Global

```{r budget_global_c_star_rmse_max}

dcant_inv_budget_sub %>%
  filter(data_source == "obs") %>% 
  group_by(c_star_rmse_max) %>% 
  summarise(rate = sum(dcant_total/duration)) %>% 
  ungroup() %>% 
  ggplot(aes(NA, rate, fill = c_star_rmse_max)) +
  geom_bar(stat="identity", position=position_dodge(), col="black") +
  scale_y_continuous() +
  labs(y = expression(Uptake~rate ~ (PgC ~ yr ^ {-1}))) +
  scale_fill_brewer(palette = "PuBuGn") +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank())

rm(dcant_inv_budget_sub)

```


# VIF max

## Column inventories

```{r column_inventory_maps_vif_max}

dcant_inv_all_bias_sub <- dcant_inv_all_bias %>%
  filter(Version_ID %in% Version_IDs[c(1,15,16)])

dcant_inv_all_bias_sub %>% 
  p_map_dcant_inv(var = "obs",
                 col = "divergent",
                 subtitle_text = "data_source: observations") +
  facet_grid(vif_max ~ .)


dcant_inv_all_bias_sub %>%
  p_map_dcant_inv(var = "dcant_bias",
                 col = "bias",
                 subtitle_text = "data_source: mod - mod_truth") +
  facet_grid(vif_max ~ .)

rm(dcant_inv_all_bias_sub)

```

## Budgets

### Regional

```{r budgets_regional_vif_max, fig.asp=1}

dcant_inv_budget_sub <- dcant_inv_budget %>%
  filter(Version_ID %in% Version_IDs[c(1,15,16)]) %>% 
  mutate(vif_max = as.factor(vif_max))

dcant_inv_budget_sub %>%
  ggplot(aes(data_source, dcant_total/duration, fill = basin_AIP)) +
  scale_fill_brewer(palette = "PuBuGn") +
  geom_col(col = "black") +
  labs(y = expression(Uptake~rate ~ (PgC ~ yr ^ {-1}))) +
  facet_grid(vif_max ~ .)

```

### Global

```{r budget_global_vif_max}

dcant_inv_budget_sub %>%
  filter(data_source == "obs") %>% 
  group_by(vif_max) %>% 
  summarise(rate = sum(dcant_total/duration)) %>% 
  ungroup() %>% 
  ggplot(aes(NA, rate, fill = vif_max)) +
  geom_bar(stat="identity", position=position_dodge(), col="black") +
  scale_y_continuous() +
  labs(y = expression(Uptake~rate ~ (PgC ~ yr ^ {-1}))) +
  scale_fill_brewer(palette = "PuBuGn") +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank())

rm(dcant_inv_budget_sub)

```


# MLR target variable

## Column inventories

```{r column_inventory_maps_MLR_target}

dcant_inv_all_bias_sub <- dcant_inv_all_bias %>%
  filter(Version_ID %in% Version_IDs[c(1,9,10)])

dcant_inv_all_bias_sub %>% 
  p_map_dcant_inv(var = "obs",
                 col = "divergent",
                 subtitle_text = "data_source: observations") +
  facet_grid(MLR_target ~ .)


dcant_inv_all_bias_sub %>%
  p_map_dcant_inv(var = "dcant_bias",
                 col = "bias",
                 subtitle_text = "data_source: mod - mod_truth") +
  facet_grid(MLR_target ~ .)

rm(dcant_inv_all_bias_sub)

```

## Budgets

### Regional

```{r budgets_regional_MLR_target, fig.asp=1}

dcant_inv_budget_sub <- dcant_inv_budget %>%
  filter(Version_ID %in% Version_IDs[c(1,9,10)]) %>% 
  mutate(MLR_target = as.factor(MLR_target))

dcant_inv_budget_sub %>%
  ggplot(aes(data_source, dcant_total/duration, fill = basin_AIP)) +
  scale_fill_brewer(palette = "PuBuGn") +
  geom_col(col = "black") +
  labs(y = expression(Uptake~rate ~ (PgC ~ yr ^ {-1}))) +
  facet_grid(MLR_target ~ .)

```

### Global

```{r budget_global_MLR_target}

dcant_inv_budget_sub %>%
  filter(data_source == "obs") %>% 
  group_by(MLR_target) %>% 
  summarise(rate = sum(dcant_total/duration)) %>% 
  ungroup() %>% 
  ggplot(aes(NA, rate, fill = MLR_target)) +
  geom_bar(stat="identity", position=position_dodge(), col="black") +
  scale_y_continuous() +
  labs(y = expression(Uptake~rate ~ (PgC ~ yr ^ {-1}))) +
  scale_fill_brewer(palette = "PuBuGn") +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank())

rm(dcant_inv_budget_sub)

```


# GLODAPv2.2021 prerelease

# Data preparation

## Load data

```{r read_files_2021_beta}

dcant_inv_all <- dcant_inv_all %>% 
  select(-era_duplication) %>% 
  mutate(GLODAPv2_version = "2020")

Version_IDs <- list.files(
  path = "/nfs/kryo/work/jenmueller/emlr_dcant/observations",
  pattern = "v_2")

# Version_IDs <- Version_IDs[1:length(Version_IDs)-1]

for (i_Version_IDs in Version_IDs) {
  # i_Version_IDs <- Version_IDs[1]
  
  print(i_Version_IDs)
  
  path_version_data     <-
  paste(path_observations,
        i_Version_IDs,
        "/data/",
        sep = "")
  
  dcant_inv <-
    read_csv(paste(path_version_data,
                   "dcant_inv.csv",
                   sep = ""))
  
  dcant_inv_mod_truth <-
    read_csv(paste(path_version_data,
                   "dcant_inv_mod_truth.csv",
                   sep = ""))
  
  dcant_inv <- bind_rows(dcant_inv, dcant_inv_mod_truth)
  
  dcant_inv <- dcant_inv %>% 
    mutate(Version_ID = i_Version_IDs)
  
  params_local <-
    read_rds(paste(path_version_data,
                   "params_local.rds",
                   sep = ""))
  
  params_local <- bind_cols(
    Version_ID = i_Version_IDs,
    MLR_basins = params_local$MLR_basins,
    MLR_target = params_local$MLR_target,
    CANYON_B_max = params_local$CANYON_B_max,
    c_star_rmse_max = params_local$c_star_rmse_max,
    vif_max = params_local$vif_max,
    GLODAPv2_version = params_local$GLODAPv2_version
  )
  
  dcant_inv <- full_join(dcant_inv, params_local)
  
  tref <- read_csv(paste(path_version_data,
                         "tref.csv",
                         sep = ""))
  
  duration <- sort(tref$median_year)[2] - sort(tref$median_year)[1]
  
  dcant_inv <- dcant_inv %>% 
    mutate(duration = duration)
  
  if (exists("dcant_inv_all")) {
    dcant_inv_all <- bind_rows(dcant_inv_all, dcant_inv)
  }
  
  if (!exists("dcant_inv_all")) {
    dcant_inv_all <- dcant_inv
  }
  
  
}

rm(dcant_inv)

```

## Format column inventories

```{r filter_standard_inventory_depth_2021_beta}

dcant_inv_all <- dcant_inv_all %>%
  filter(inv_depth == params_global$inventory_depth_standard)

```

```{r calc_column_inventories_2021_beta}

dcant_inv_all_bias <- dcant_inv_all %>%
  select(lat, lon, data_source, dcant_inv, Version_ID, 
         GLODAPv2_version, MLR_basins,
         MLR_target, CANYON_B_max, vif_max, c_star_rmse_max) %>%
  pivot_wider(names_from = data_source,
              values_from = dcant_inv) %>%
  mutate(dcant_bias = mod - mod_truth) %>%
  drop_na()

```

## Calculate budgets

```{r calc_budgets_2021_beta}

dcant_inv_budget <- dcant_inv_all %>% 
  mutate(surface_area = earth_surf(lat, lon),
         dcant_inv_grid = dcant_inv*surface_area,
         dcant_pos_inv_grid = dcant_pos_inv*surface_area) %>% 
  group_by(basin_AIP, data_source,
           Version_ID, MLR_basins,
           GLODAPv2_version, duration,
           MLR_target, CANYON_B_max, vif_max, c_star_rmse_max) %>% 
  summarise(dcant_total = sum(dcant_inv_grid)*12*1e-15,
            dcant_total = round(dcant_total,1),
            dcant_pos_total = sum(dcant_pos_inv_grid)*12*1e-15,
            dcant_pos_total = round(dcant_pos_total,1)) %>% 
  ungroup()

dcant_inv_budget %>%
  group_by(data_source, Version_ID) %>%
  summarise(dcant_total = sum(dcant_total),
            dcant_total = round(dcant_total,1),
            dcant_pos_total = sum(dcant_pos_total),
            dcant_pos_total = round(dcant_pos_total,1),
            dcant_total_rate = dcant_total / duration,
            dcant_pos_total_rate = dcant_pos_total / duration) %>%
  ungroup()

```


# MLR regions and era overlap

## Column inventories

```{r column_inventory_maps_regions_eras_2021_beta, fig.asp=1}

dcant_inv_all_bias_sub <- dcant_inv_all_bias %>%
  filter(Version_ID %in% c("v_101", "v_102", "v_103", "v_104",
                           "v_201", "v_202", "v_203", "v_204"))


dcant_inv_all_bias_sub %>% 
  p_map_dcant_inv(var = "obs",
                 col = "divergent",
                 subtitle_text = "data_source: observations") +
  facet_grid(MLR_basins ~ GLODAPv2_version)


dcant_inv_all_bias_sub %>%
  p_map_dcant_inv(var = "dcant_bias",
                 col = "divergent",
                 subtitle_text = "data_source: mod - mod_truth") +
  facet_grid(MLR_basins ~ GLODAPv2_version)


```

## Budgets

### Regional

```{r budgets_regional_regions_eras_2021_beta, fig.asp=1}

dcant_inv_budget_sub <- dcant_inv_budget %>%
  filter(Version_ID %in% unique(dcant_inv_all_bias_sub$Version_ID))

dcant_inv_budget_sub %>%
  ggplot(aes(data_source, dcant_total/duration, fill = basin_AIP)) +
  scale_fill_brewer(palette = "PuBuGn") +
  geom_col(col = "black") +
  labs(y = expression(Uptake~rate ~ (PgC ~ yr ^ {-1}))) +
  facet_grid(MLR_basins ~ GLODAPv2_version)

```

### Global

```{r budget_global_regions_eras_2021_beta}

dcant_inv_budget_sub %>%
  filter(data_source == "obs") %>% 
  group_by(MLR_basins, GLODAPv2_version) %>% 
  summarise(rate = sum(dcant_total/duration)) %>% 
  ungroup() %>% 
  ggplot(aes(GLODAPv2_version, rate, fill=MLR_basins)) +
  geom_bar(stat="identity", position=position_dodge(), col="black") +
  scale_y_continuous() +
  labs(y = expression(Uptake~rate ~ (PgC ~ yr ^ {-1})),
       x = "") +
  scale_fill_brewer(palette = "PuBuGn")

rm(dcant_inv_budget_sub)

```




