```{r r join_data_and_meta_data, include=FALSE}


dcant_inv_all <- full_join(dcant_inv_all,
                           params_local_all)

dcant_inv_bias_all <- full_join(dcant_inv_bias_all,
                                params_local_all)


dcant_budget_lat_grid_all <- full_join(dcant_budget_lat_grid_all,
                                       params_local_all)

dcant_budget_lon_grid_all <- full_join(dcant_budget_lon_grid_all,
                                       params_local_all)

```


```{r adapt_format}

dcant_budget_lat_grid_all <- dcant_budget_lat_grid_all %>%
  pivot_wider(names_from = estimate,
              values_from = value) %>%
  filter(period != "1994 - 2014",
         method == "total")

dcant_budget_lon_grid_all <- dcant_budget_lon_grid_all %>%
  pivot_wider(names_from = estimate,
              values_from = value) %>%
  filter(period != "1994 - 2014",
         method == "total")


```


# Uncertainty limit

```{r define_sd_uncertainty_limit}

sd_uncertainty_limit <- 4

```



# Individual cases

## Absoulte values

```{r cases_absolute, fig.asp=1.2}

dcant_inv_all %>%
  filter(#data_source %in% c("mod", "obs"),
         period != "1994 - 2014") %>%
  group_by(data_source) %>%
  group_split() %>%
  # head(1) %>%
  map(
    ~ p_map_cant_inv(df = .x,
                     var = "dcant",
                     subtitle_text = paste("data_source:",
                                           unique(.x$data_source))) +
      facet_grid({{config}} ~ period) +
      theme(axis.text = element_blank(),
            axis.ticks = element_blank())
  )

dcant_inv_all %>%
  filter(data_source %in% c("mod", "obs"),
         period == "1994 - 2014") %>%
  p_map_cant_inv(
    var = "dcant",
    subtitle_text = "1994 - 2014",
    breaks = c(-Inf, seq(0, 16, 2), Inf) * 2
  ) +
  facet_grid({{config}} ~ data_source) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())


p_dcant_inv_all_1994_2004 <-
  dcant_inv_all %>%
  filter(data_source %in% c("obs"),
         period == "1994 - 2004") %>%
  mutate(period = recode(period,
                         "1994 - 2004" = "Observation-based")) %>%
  p_map_cant_inv(var = "dcant",
                 title_text = "1994 - 2004") +
  facet_grid({{config}} ~ period) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    strip.background.y = element_blank(),
    strip.text.y = element_blank(),
    legend.position = "left"
  )


```

## Biases

```{r cases_bias, fig.asp=1.2}


dcant_inv_bias_all %>%
  filter(period != "1994 - 2014") %>%
  p_map_cant_inv(var = "dcant_bias",
                 col = "bias",
                 subtitle_text = "data_source: mod - mod_truth") +
  facet_grid({{config}} ~ period)

dcant_inv_bias_all %>%
  filter(period == "1994 - 2014") %>%
  p_map_cant_inv(var = "dcant_bias",
                 col = "bias",
                 subtitle_text = "data_source: mod - mod_truth") +
  facet_grid({{config}} ~ period)


p_dcant_inv_bias_all_1994_2004 <-
  dcant_inv_bias_all %>%
  filter(period == "1994 - 2004") %>%
  mutate(period = recode(period,
                         "1994 - 2004" = "Model-based")) %>%
  p_map_cant_inv(var = "dcant_bias",
                 col = "bias",
                 title_text = "Models") +
  facet_grid({{config}} ~ period) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    plot.title = element_blank()
  )


```

```{r composed_absolute_and_bias_figure, fig.asp=1.2}

p_dcant_1994_2004 <- 
p_dcant_inv_all_1994_2004 +
  p_dcant_inv_bias_all_1994_2004

p_dcant_1994_2004

# ggsave(plot = p_dcant_1994_2004,
#        path = "output/other",
#        filename = "inv_dcant_1994_2004_abs_bias.png",
#        height = 8,
#        width = 8)

```

## Decadal offset

```{r decadal_dcant_offset_column_inventory_map, fig.asp=1.2}

dcant_inv_all %>%
  filter(period != "1994 - 2014") %>%
  select(data_source, lon, lat, {{config}}, period, dcant_pos) %>% 
  pivot_wider(values_from = dcant_pos,
              names_from = period) %>% 
  mutate(dcant_pos_bias = `2004 - 2014`- `1994 - 2004`) %>% 
  group_by(data_source) %>%
  group_split() %>%
  # head(1) %>%
  map(
    ~ p_map_cant_inv(df = .x,
                     var = "dcant_pos_bias",
                     col = "bias",
                     subtitle_text = paste("data_source:",
                                           unique(.x$data_source))) +
      facet_grid({{config}} ~ .) +
      theme(axis.text = element_blank(),
            axis.ticks = element_blank())
  )



```


```{r decadal_bias_offset_column_inventory_map, fig.asp=1.2}

dcant_inv_bias_all %>%
  filter(period != "1994 - 2014") %>%
  select(lon, lat, {{config}}, period, dcant_pos_bias) %>%
  pivot_wider(values_from = dcant_pos_bias,
              names_from = period) %>%
  mutate(dcant_pos_bias = `2004 - 2014` - `1994 - 2004`) %>%
  p_map_cant_inv(var = "dcant_pos_bias",
                 col = "bias") +
  facet_grid({{config}} ~ .) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())


```



### Density distribution

```{r bias_density_distribution, fig.asp=1}

dcant_inv_bias_all %>%
  filter(abs(dcant_bias) < 10) %>% 
  ggplot(aes(dcant_bias, col = {{config}})) +
  scale_color_brewer(palette = "Dark2") +
  geom_vline(xintercept = 0) +
  geom_density() +
  facet_grid(period ~.)

```


## Lat grid budgets

```{r lat_grid_budget_all, fig.asp=2}

dcant_budget_lat_grid_all %>%
  group_split(data_source) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(lat_grid, dcant, fill = {{config}})) +
      geom_hline(yintercept = 0) +
      geom_col(position = "dodge") +
      coord_flip() +
      scale_fill_brewer(palette = "Dark2") +
      labs(title = paste("data_source:", unique(.x$data_source))) +
      facet_grid(basin_AIP ~ period)
  )

```

## Lon grid budgets

```{r lon_grid_budget_all}

dcant_budget_lon_grid_all %>%
  group_split(data_source, period) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x,
             aes(lon_grid, dcant, fill = {{config}})) +
      geom_col(position = "dodge") +
      scale_fill_brewer(palette = "Dark2") +
      labs(title = paste(
        "data_source:",
        unique(.x$data_source),
        "| period:",
        unique(.x$period)
      )) +
      facet_grid(basin_AIP ~ .)
  )

```


# Ensemble


```{r calc_ensemble_metrics}

dcant_inv_ensemble <- dcant_inv_all %>% 
  filter(data_source %in% c("mod", "obs")) %>% 
  group_by(lat, lon, data_source, period) %>% 
  summarise(dcant_mean = mean(dcant),
            dcant_sd = sd(dcant),
            dcant_range = max(dcant)- min(dcant)) %>% 
  ungroup()

dcant_lat_grid_ensemble <- dcant_budget_lat_grid_all %>% 
  filter(data_source %in% c("mod", "obs")) %>% 
  group_by(lat_grid, data_source, period, {{config}}) %>% 
  summarise(dcant = sum(dcant, na.rm = TRUE)) %>% 
  ungroup() %>% 
  group_by(lat_grid, data_source, period) %>% 
  summarise(dcant_mean = mean(dcant),
            dcant_sd = sd(dcant),
            dcant_max = max(dcant),
            dcant_min = min(dcant)) %>% 
  ungroup()

dcant_lon_grid_ensemble <- dcant_budget_lon_grid_all %>% 
  filter(data_source %in% c("mod", "obs")) %>% 
  group_by(lon_grid, data_source, period, {{config}}) %>% 
  summarise(dcant = sum(dcant, na.rm = TRUE)) %>% 
  ungroup() %>% 
  group_by(lon_grid, data_source, period) %>% 
  summarise(dcant_mean = mean(dcant),
            dcant_sd = sd(dcant),
            dcant_max = max(dcant),
            dcant_min = min(dcant)) %>% 
  ungroup()


```

## Mean

```{r ensemble_mean}

p_map_cant_inv(
  df = dcant_inv_ensemble %>%
    filter(period != "1994 - 2014",
           data_source == "obs"),
  var = "dcant_mean",
  subtitle_text = paste("Ensemble mean")
) +
  facet_grid(period ~ .) +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank())

```

```{r ensemble_mean_two_decades, fig.asp=1}

p_map_cant_inv(
  df = dcant_inv_ensemble %>%
    filter(data_source == "obs",
           period == "1994 - 2014"),
  var = "dcant_mean",
  title_text = paste("Ensemble mean"),
  breaks = c(-Inf, seq(0, 28, 4), Inf)
) +
  facet_grid(period ~ .) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "left"
  )

```

## Mean bias

```{r ensemble_mean_bias}

dcant_inv_ensemble_bias <- full_join(
  dcant_inv_ensemble %>%
    filter(data_source == "mod") %>%
    select(lat, lon, period, dcant_mean, dcant_sd),
  dcant_inv_all %>%
    filter(data_source == "mod_truth",
           {{config}} == unique(dcant_inv_all${{config}})[1]) %>%
    select(lat, lon, period, dcant)
)

dcant_inv_ensemble_bias <- dcant_inv_ensemble_bias %>%
  mutate(dcant_mean_bias = dcant_mean - dcant)

dcant_inv_ensemble_bias %>%
  filter(period != "1994 - 2014") %>%
  p_map_cant_inv(var = "dcant_mean_bias",
                 col = "bias",
                 subtitle_text = "Ensemble mean - mod_truth") +
  facet_grid(period ~ .)

```

### Density distribution

```{r bias_density_distribution_ensemble, fig.asp=1}

dcant_inv_bias_all %>%
  ggplot() +
  scale_color_manual(values = c("red", "grey")) +
  geom_vline(xintercept = 0) +
  geom_density(aes(dcant_bias, group = {{config}}, col = "Individual")) +
  geom_density(data = dcant_inv_ensemble_bias,
               aes(dcant_mean_bias, col = "Ensemble")) +
  facet_grid(period ~.) +
  coord_cartesian(xlim = c(-10, 10))

```


## Mean lat grid budgets

```{r lat_grid_budget_ensemble, fig.asp=1}

dcant_lat_grid_ensemble %>%
  ggplot(aes(lat_grid, dcant_mean)) +
  geom_hline(yintercept = 0) +
  geom_col(position = "dodge",
           fill = "grey80",
           col = "grey20") +
  geom_errorbar(aes(
    ymin = dcant_min,
    ymax = dcant_max
  ),
  col = "grey20",
  width = 0) +
  scale_color_brewer(palette = "Set1") +
  coord_flip() +
  scale_fill_brewer(palette = "Dark2") +
  facet_grid(data_source ~ period)

```

## Mean lon grid budgets

```{r lon_grid_budget_ensemble}

dcant_lon_grid_ensemble %>%
  ggplot(aes(lon_grid, dcant_mean)) +
  geom_col(position = "dodge",
           fill = "grey80",
           col = "grey20") +
  geom_errorbar(aes(
    ymin = dcant_min,
    ymax = dcant_max
  ),
  col = "grey20",
  width = 0) +
  facet_grid(data_source ~ period)


```


## Standard deviation

```{r ensemble_sd, fig.asp=0.5}

p_map_cant_inv(
  df = dcant_inv_ensemble,
  var = "dcant_sd",
  breaks = c(seq(0, 4, 0.4), Inf),
  subtitle_text = paste("Ensemble SD")
) +
  facet_grid(period ~ data_source)

```

## SD as uncertainty

```{r ensemble_sd_uncertainty, fig.asp=0.5}

uncertainty_grid <- dcant_inv_ensemble %>%
  filter(dcant_sd > sd_uncertainty_limit) %>%
  distinct(lon, lat, data_source, period)

uncertainty_grid <- m_grid_horizontal_coarse(uncertainty_grid) %>%
  distinct(lon_grid, lat_grid, data_source, period)

map +
  geom_point(data =
               uncertainty_grid,
             aes(lon_grid, lat_grid),
             shape = 3) +
  facet_grid(period ~ data_source)

```


## SD vs abs bias

### 2D bin

```{r ensemble_sd_vs_bias}

dcant_inv_ensemble_bias %>% 
  ggplot(aes(abs(dcant_mean_bias), dcant_sd)) +
  geom_bin2d() +
  scale_fill_viridis_c() +
  facet_grid(. ~ period)



```

### Density distributions

```{r density_distributions}

legend_title = expression(Delta * C[ant]~(mol~m^2))

dcant_density_distribution_bias <- dcant_inv_ensemble_bias %>% 
  select(dcant_mean_bias, period) %>% 
  pivot_longer(dcant_mean_bias,
               names_to = "estimate",
               values_to = "value") %>% 
  mutate(data_source = "mod")

dcant_density_distribution <- dcant_inv_ensemble %>% 
  select(dcant_mean, dcant_sd, period, data_source) %>% 
  pivot_longer(dcant_mean:dcant_sd,
               names_to = "estimate",
               values_to = "value")

dcant_density_distribution <- bind_rows(dcant_density_distribution,
                                        dcant_density_distribution_bias) %>%
  mutate(period = factor(period, c("1994 - 2004", "2004 - 2014", "1994 - 2014")))

p_ensemble_distribution <-
  dcant_density_distribution %>%
  filter(period != "1994 - 2014") %>% 
  mutate(
    data_source = recode(data_source,
                         mod = "Model",
                         obs = "Observations"),
    estimate = recode(
      estimate,
      dcant_mean = "Mean",
      dcant_mean_bias = "Bias",
      dcant_sd = "SD"
    )
  ) %>%
  ggplot(aes(value, fill = estimate, col = estimate)) +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  geom_density(alpha = 0.2) +
  facet_grid(period ~ data_source) +
  labs(title = "Density distributions of column inventory ensemble properties",
       x = legend_title,
       y = "Density of 1x1° grid cells") +
  coord_cartesian(ylim = c(0, 0.5),
                  xlim = c(-5, 32))

# ggsave(plot = p_ensemble_distribution,
#        path = "output/other",
#        filename = "inv_ensemble_distribution.png",
#        height = 4,
#        width = 8)


p_ensemble_distribution <-
  dcant_density_distribution %>%
  filter(data_source == "obs") %>%
  mutate(
    data_source = recode(data_source,
                         mod = "Model",
                         obs = "Observations"),
    estimate = recode(
      estimate,
      dcant_mean = "Mean",
      dcant_mean_bias = "Bias",
      dcant_sd = "SD"
    )
  ) %>%
  ggplot(aes(value, fill = estimate, col = estimate)) +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  geom_density(alpha = 0.2) +
  facet_grid(period ~ data_source) +
  labs(title = "Density distributions of column inventory ensemble properties",
       x = legend_title,
       y = "Density of 1x1° grid cells") +
  coord_cartesian(ylim = c(0, 0.5),
                  xlim = c(-5, 32))

p_ensemble_distribution

# ggsave(plot = p_ensemble_distribution,
#        path = "output/other",
#        filename = "inv_ensemble_distribution_all.png",
#        height = 6,
#        width = 5)

```



## Range

```{r ensemble_range, fig.asp=0.5}

p_map_cant_inv(
  df = dcant_inv_ensemble,
  var = "dcant_range",
  breaks = c(seq(0,8,0.8), Inf),
  subtitle_text = paste("Ensemble range")
) +
  facet_grid(period ~ data_source)



```

## Composed figure

```{r composed_figure}

uncertainty_grid <- uncertainty_grid %>% 
  filter(data_source == "obs")

p_map_94 <- p_map_cant_inv(
  df = dcant_inv_ensemble %>%
    filter(period == "1994 - 2004",
           data_source == "obs"),
  var = "dcant_mean",
  subtitle_text = NULL,
  title_text = NULL
) +
  geom_point(
    data =
      uncertainty_grid %>% filter(period == "1994 - 2004"),
    aes(lon_grid, lat_grid),
    shape = 3,
    col = "red"
  ) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "left"
  ) +
  facet_grid(period ~ .,
             switch = "y")

p_map_04 <- p_map_cant_inv(
  df = dcant_inv_ensemble %>%
    filter(period == "2004 - 2014",
           data_source == "obs"),
  var = "dcant_mean",
  subtitle_text = NULL,
  title_text = NULL
) +
  geom_point(
    data =
      uncertainty_grid %>% filter(period == "2004 - 2014"),
    aes(lon_grid, lat_grid),
    shape = 3,
    col = "red"
  ) +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "left"
  ) +
  facet_grid(period ~ .,
             switch = "y")

p_lon_94 <- dcant_lon_grid_ensemble %>%
  filter(data_source == "obs",
         period == "1994 - 2004") %>%
  ggplot(aes(lon_grid, dcant_mean)) +
  geom_col(position = "dodge",
           fill = "grey80",
           col = "grey20") +
  geom_errorbar(aes(ymin = dcant_min,
                    ymax = dcant_max),
                col = "grey20",
                width = 0) +
  scale_y_continuous(
    limits = c(0, 1),
    expand = c(0, 0),
    name = expression(Delta * C[ant] ~ (PgC)),
    position = "right"
  ) +
  scale_x_continuous(name = "Longitude (°E)",
                     limits = c(20, 380),
                     expand = c(0, 0)) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
# coord_fixed(ratio = 100)

p_lon_04 <- dcant_lon_grid_ensemble %>%
  filter(data_source == "obs",
         period == "2004 - 2014") %>% 
  ggplot(aes(lon_grid, dcant_mean)) +
  geom_col(position = "dodge",
           fill = "grey80",
           col = "grey20") +
  geom_errorbar(aes(
    ymin = dcant_min,
    ymax = dcant_max
  ),
  col = "grey20",
  width = 0) +
  scale_y_continuous(limits = c(0,1), expand = c(0,0),
                     name = expression(Delta*C[ant]~(PgC)),
                     position = "right") +
  scale_x_continuous(name = "Longitude (°E)",
                     limits = c(20,380), expand = c(0,0)) +
  theme(title = element_blank())
  # coord_fixed(ratio = 100)

p_lat_94 <- dcant_lat_grid_ensemble %>%
  filter(data_source == "obs",
         period == "1994 - 2004") %>%
  ggplot(aes(dcant_mean, lat_grid)) +
  geom_col(
    position = "dodge",
    fill = "grey80",
    col = "grey20",
    orientation = "y"
  ) +
  geom_errorbar(aes(xmin = dcant_min,
                    xmax = dcant_max),
                col = "grey20",
                width = 0) +
  scale_x_continuous(
    limits = c(0, 2.8),
    expand = c(0, 0),
    name = expression(Delta * C[ant] ~ (PgC))
  ) +
  scale_y_continuous(
    name = "Latitude (°N)",
    limits = c(-80, 65),
    expand = c(0, 0),
    position = "right"
  ) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ) 
  # coord_fixed(ratio = 5e-2)

p_lat_04 <- dcant_lat_grid_ensemble %>%
  filter(data_source == "obs",
         period == "2004 - 2014") %>%
  ggplot(aes(dcant_mean, lat_grid)) +
  geom_col(position = "dodge",
           fill = "grey80",
           col = "grey20",
           orientation = "y") +
  geom_errorbar(aes(xmin = dcant_min,
                    xmax = dcant_max),
                col = "grey20",
                width = 0) +
  scale_x_continuous(
    limits = c(0, 2.8),
    expand = c(0, 0),
    name = expression(Delta * C[ant] ~ (PgC))
  ) +
  scale_y_continuous(name = "Latitude (°N)",
                     limits = c(-80, 65),
                     expand = c(0, 0),
                     position = "right")

t_94 <- grid::textGrob("1994 -\n2004")
t_04 <- grid::textGrob("2004 -\n2014")


layout <- "
BBB#
AAAD
AAAD
CCCE
CCCE
GGG#
"

wrap_plots(B = p_lon_94, A = p_map_94, D = p_lat_94,
           G = p_lon_04, C = p_map_04, E = p_lat_04,
           design = layout,
           guides = "collect")&
  theme(legend.position = "left")


# ggsave("output/publication/Fig_column_inventories.png",
#        width=11,
#        height=9.27)


```




# Cases vs ensemble

## Offset from mean

```{r ensemble_deviation_from_mean}

dcant_inv_all <- full_join(dcant_inv_all,
                           dcant_inv_ensemble)

dcant_inv_all <- dcant_inv_all %>% 
  mutate(dcant_offset = dcant - dcant_mean)

dcant_inv_all %>%
  filter(data_source %in% c("mod", "obs")) %>%
  group_by(period) %>%
  group_split() %>%
  # head(1) %>%
  map(
    ~ p_map_cant_inv(df = .x,
                     var = "dcant_offset",
                     col = "bias",
                     subtitle_text = paste("period:",
                                           unique(.x$period))) +
      facet_grid({{config}} ~ data_source)
  )


```



