---
title: "Anomalous changes"
author: "Jens Daniel Müller"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  workflowr::wflow_html:
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

```{r parent, child = "/nfs/kryo/work/jenmueller/emlr_cant/utilities/setup.Rmd"}
# this chunk runs the code stored in setup.Rmd
# if required, please refer to instructions given here:
# https://jdblischak.github.io/workflowr/articles/wflow-07-common-code.html
```

```{r define_path, include = FALSE}

# only path_observations needs to be changed to model
path_observations <-
  paste(path_root, "/observations/", sep = "")

path_preprocessing    <-
  paste(path_observations, "preprocessing/", sep = "")


path_preprocessing_model    <-
  paste(path_root, "/model/preprocessing/", sep = "")

```

```{r load_libraries_specific, include = FALSE}
library(marelac)
library(ggforce)
```

# Data preparation

```{r read_version_IDs}

Version_IDs <- list.files(
  path = "/nfs/kryo/work/jenmueller/emlr_cant/observations",
  pattern = "v_2")

# Version_IDs <- Version_IDs[1:4]

```

## Read files

Following sensitivity cases were considered:

```{r read_cant_files}

for (i_Version_IDs in Version_IDs) {
  # i_Version_IDs <- Version_IDs[1]
  
  print(i_Version_IDs)
  
  path_version_data     <-
  paste(path_observations,
        i_Version_IDs,
        "/data/",
        sep = "")
  
  # read inventory file
  cant_inv <-
    read_csv(paste(path_version_data,
                   "anom_cant_mod_inv_bias.csv",
                   sep = ""))
  
  cant_inv <- cant_inv %>% 
    mutate(Version_ID = i_Version_IDs)
  
  # read section file
  cant_section <-
    read_csv(paste(path_version_data,
                   "anom_cant_mod_section_bias.csv",
                   sep = ""))
  
  cant_section <- cant_section %>% 
    mutate(Version_ID = i_Version_IDs)
  
  # read budget file
  cant_budget <-
    read_csv(paste(path_version_data,
                   "anom_cant_mod_inv_budget.csv",
                   sep = ""))
  
  cant_budget <- cant_budget %>% 
    mutate(Version_ID = i_Version_IDs)
  
  # read local parametrization files
  params_local <-
    read_rds(paste(path_version_data,
                   "params_local.rds",
                   sep = ""))
  
  params_local <- bind_cols(
    Version_ID = i_Version_IDs,
    MLR_basins = params_local$MLR_basins,
    rarefication = params_local$rarefication,
    rarefication_threshold = params_local$rarefication_threshold,
    vif_max = params_local$vif_max
  )
  
  cant_inv <- full_join(cant_inv, params_local)
  cant_section <- full_join(cant_section, params_local)
  cant_budget <- full_join(cant_budget, params_local)

  # read tref file
  tref <- read_csv(paste(path_version_data,
                         "tref.csv",
                         sep = ""))
  
  duration <- sort(tref$median_year)[2] - sort(tref$median_year)[1]
  eras <- paste0(sort(tref$median_year)[1],"-", sort(tref$median_year)[2])
  
  cant_inv <- cant_inv %>% 
    mutate(duration = duration,
           eras = eras)
  
  cant_section <- cant_section %>%
    mutate(duration = duration,
           eras = eras)
  
  cant_budget <- cant_budget %>%
    mutate(duration = duration,
           eras = eras)
  
  # bind inv files
  if (exists("cant_inv_all")) {
    cant_inv_all <- bind_rows(cant_inv_all, cant_inv)
  }
  
  if (!exists("cant_inv_all")) {
    cant_inv_all <- cant_inv
  }

  # bind section files
  if (exists("cant_section_all")) {
    cant_section_all <- bind_rows(cant_section_all, cant_section)
  }
  
  if (!exists("cant_section_all")) {
    cant_section_all <- cant_section
  }
  
  # bind budget files
  if (exists("cant_budget_all")) {
    cant_budget_all <- bind_rows(cant_budget_all, cant_budget)
  }
  
  if (!exists("cant_budget_all")) {
    cant_budget_all <- cant_budget
  }
  
  
}

rm(cant_inv, cant_section, cant_budget)

```


### Format

```{r format_cant_files}

cant_inv_all_long <- cant_inv_all %>%
  pivot_longer(emlr_anom:nss_projection,
               names_to = "estimate",
               values_to = "delta_cant_offset")


cant_section_all_long <- cant_section_all %>%
  pivot_longer(emlr_anom:nss_projection,
               names_to = "estimate",
               values_to = "delta_cant_offset")

# unique(cant_inv_all$Version_ID)

```


# Eras vs basin split

Here we compare the impact of the basin separation approach on the anomaly detection, separately for each compared eras.

```{r subset_era_basin_split}

cant_inv_all_sub <- cant_inv_all %>% 
    filter(rarefication == "none")

cant_inv_all_long_sub <- cant_inv_all_long %>% 
    filter(rarefication == "none")

cant_section_all_sub <- cant_section_all %>% 
    filter(rarefication == "none")

cant_section_all_long_sub <- cant_section_all_long %>% 
    filter(rarefication == "none")

cant_budget_all_sub <- cant_budget_all %>% 
    filter(rarefication == "none")

```

## Column inventories

### Maps

```{r column_inventory_maps_era_basin_split}

cant_inv_all_long_sub %>%
  group_split(eras) %>%
  # head(1) %>%
  map(
    ~ p_map_cant_inv_offset(
      df = .x,
      var = "delta_cant_offset",
      # col = "divergent",
      subtitle_text = paste("Eras:", unique(.x$eras)),
      breaks = seq(-8,8,1)
    ) +
      facet_grid(MLR_basins ~ estimate) +
      theme(axis.text = element_blank(),
            axis.ticks = element_blank())
  )


```


### Bias correlation

```{r inv_bias_plots_era_basin_split, fig.asp=1}

axis_limit <- 10

cant_inv_all_sub %>%
  group_split(eras) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x, aes(emlr_bias, emlr_anom)) +
      geom_hline(yintercept = 0) +
      geom_vline(xintercept = 0) +
      geom_bin2d(binwidth = 1) +
      scale_fill_viridis_c(trans = "log10") +
      coord_equal(
        xlim = c(-axis_limit, axis_limit),
        ylim = c(-axis_limit, axis_limit)
      ) +
      labs(title = paste("Eras:", unique(.x$eras))) +
      facet_grid(MLR_basins ~ basin_AIP)
  )


cant_inv_all_sub %>%
  group_split(eras) %>%
  # head(1) %>% 
  map(
    ~ ggplot(data = .x, aes(emlr_anom, nss_mod_truth)) +
      geom_hline(yintercept = 0) +
      geom_vline(xintercept = 0) +
      geom_bin2d(binwidth = 1) +
      scale_fill_viridis_c(trans = "log10") +
      coord_equal(
        xlim = c(-axis_limit, axis_limit),
        ylim = c(-axis_limit, axis_limit)
      ) +
      labs(title = paste("Eras:", unique(.x$eras))) +
      facet_grid(MLR_basins ~ basin_AIP)
  )

cant_inv_all_sub %>%
  group_split(eras) %>%
  # head(1) %>% 
  map(
    ~ ggplot(data = .x, aes(nss_mod_truth, nss_projection)) +
      geom_hline(yintercept = 0) +
      geom_vline(xintercept = 0) +
      geom_bin2d(binwidth = 1) +
      scale_fill_viridis_c(trans = "log10") +
      coord_equal(
        xlim = c(-axis_limit, axis_limit),
        ylim = c(-axis_limit, axis_limit)
      ) +
      labs(title = paste("Eras:", unique(.x$eras))) +
      facet_grid(MLR_basins ~ basin_AIP)
  )

```


## Zonal sections

### Sections

```{r cant_sections_era_basin_split, fig.asp=0.5, eval=FALSE}

cant_section_all_long_sub %>%
  group_split(eras, MLR_basins, estimate) %>%
  # head(1) %>%
  map(
    ~ p_section_zonal(
      df = .x,
      var = "delta_cant_offset",
      plot_slabs = "n",
      col = "divergent",
      subtitle_text = paste("Eras:", unique(.x$eras),
                            "| MLR Basins:", unique(.x$MLR_basins),
                            "| Estimate:", unique(.x$estimate)),
      breaks = c(-Inf,seq(-8,8,1), Inf)
    )
  )


```


```{r cant_sections_era_basin_split_facet}

legend_title <- expression(atop(Delta * C[ant, pos],
                               (mu * mol ~ kg ^ {
                                 -1
                               })))

cant_section_all_long_sub %>%
  group_split(eras, basin_AIP) %>% 
  # head(1) %>%
  map(
    ~ ggplot(data = .x) +
      geom_contour_filled(aes(lat, depth, z = delta_cant_offset),
                          breaks =c(-Inf,seq(-8,8,1), Inf)) +
      scale_fill_scico_d(
        palette = "vik",
        drop = FALSE,
        name = legend_title
      ) +
      guides(fill = guide_colorsteps(barheight = unit(8, "cm"),
                                     show.limits = TRUE)) +
      # scale_y_sqrt() +
      scale_y_continuous(trans = trans_reverser("sqrt"),
                         breaks = c(100,500,seq(1000,5000,1000))) +
      #scale_y_reverse() +
      scale_x_continuous(breaks = seq(-100, 100, 40),
                         limits = c(-85, 85)) +
      facet_grid(MLR_basins ~ estimate) +
      coord_cartesian(expand = 0) +
      labs(title = paste("Eras", unique(.x$eras),
                         "| Basin:", unique(.x$basin_AIP)))
  )


```

### Bias correlation

```{r section_bias_plots_era_basin_split, fig.asp=1}

axis_limit <- 20

cant_section_all_sub %>%
  group_split(eras) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x, aes(emlr_bias, emlr_anom)) +
      geom_hline(yintercept = 0) +
      geom_vline(xintercept = 0) +
      geom_bin2d(binwidth = 1) +
      scale_fill_viridis_c(trans = "log10") +
      coord_equal(
        xlim = c(-axis_limit, axis_limit),
        ylim = c(-axis_limit, axis_limit)
      ) +
      labs(title = paste("Eras:", unique(.x$eras))) +
      facet_grid(MLR_basins ~ basin_AIP)
  )


cant_section_all_sub %>%
  group_split(eras) %>%
  # head(1) %>% 
  map(
    ~ ggplot(data = .x, aes(emlr_anom, nss_mod_truth)) +
      geom_hline(yintercept = 0) +
      geom_vline(xintercept = 0) +
      geom_bin2d(binwidth = 1) +
      scale_fill_viridis_c(trans = "log10") +
      coord_equal(
        xlim = c(-axis_limit, axis_limit),
        ylim = c(-axis_limit, axis_limit)
      ) +
      labs(title = paste("Eras:", unique(.x$eras))) +
      facet_grid(MLR_basins ~ basin_AIP)
  )

cant_section_all_sub %>%
  group_split(eras) %>%
  # head(1) %>% 
  map(
    ~ ggplot(data = .x, aes(nss_mod_truth, nss_projection)) +
      geom_hline(yintercept = 0) +
      geom_vline(xintercept = 0) +
      geom_bin2d(binwidth = 1) +
      scale_fill_viridis_c(trans = "log10") +
      coord_equal(
        xlim = c(-axis_limit, axis_limit),
        ylim = c(-axis_limit, axis_limit)
      ) +
      labs(title = paste("Eras:", unique(.x$eras))) +
      facet_grid(MLR_basins ~ basin_AIP)
  )

```



## Budgets

### Regional

```{r budgets_regional_era_basin_split, fig.asp=1}

cant_budget_all_sub %>%
  filter(inv_depth == params_global$inventory_depth_standard) %>% 
  ggplot(aes(estimate, cant_pos_total, fill = basin_AIP)) +
  scale_fill_brewer(palette = "PuBuGn") +
  geom_col(col = "black") +
  facet_grid(MLR_basins ~ eras) +
  theme(axis.text.x = element_text(angle = 90))

```

### Era additive

```{r budgets_regional_era_basin_split_era_additive, fig.asp=1}

cant_budget_all_sub_wide <- cant_budget_all_sub %>%
  filter(inv_depth == params_global$inventory_depth_standard) %>% 
  select(-c(duration, Version_ID)) %>% 
  pivot_wider(names_from = eras,
              values_from = cant_pos_total)

cant_budget_all_sub_wide <- cant_budget_all_sub_wide %>%
  mutate(`1994-2014 (sum)` = `1994-2006` + `2006-2014`)

cant_budget_all_sub_long <- cant_budget_all_sub_wide %>%
  select(-c(`1994-2006`, `2006-2014`)) %>% 
  pivot_longer(c(`1994-2014`, `1994-2014 (sum)`),
               names_to = "eras",
               values_to = "cant_pos_total")

cant_budget_all_sub_long %>% 
  ggplot(aes(eras, cant_pos_total, fill = basin_AIP)) +
  scale_fill_brewer(palette = "PuBuGn") +
  geom_col(col = "black") +
  facet_grid(MLR_basins ~ estimate) +
  theme(axis.text.x = element_text(angle = 90))

```





# Rarefication and VIF

```{r subset_rarefication_vif}

cant_inv_all_sub <- cant_inv_all %>% 
    filter(rarefication == "coarse_grid")

cant_inv_all_long_sub <- cant_inv_all_long %>% 
    filter(rarefication == "coarse_grid")

cant_section_all_sub <- cant_section_all %>% 
    filter(rarefication == "coarse_grid")

cant_section_all_long_sub <- cant_section_all_long %>% 
    filter(rarefication == "coarse_grid")

cant_budget_all_sub <- cant_budget_all %>% 
    filter(rarefication == "coarse_grid")

```

## Column inventories

### Maps rarefication

```{r column_inventory_maps_rarefication}

cant_inv_all_long_sub %>%
  filter(vif_max == 1000) %>% 
  group_split(eras) %>%
  # head(1) %>%
  map(
    ~ p_map_cant_inv_offset(
      df = .x,
      var = "delta_cant_offset",
      # col = "divergent",
      subtitle_text = paste("Rarefication test | Eras:", unique(.x$eras)),
      breaks = seq(-8,8,1)
    ) +
      facet_grid(rarefication_threshold ~ estimate) +
      theme(axis.text = element_blank(),
            axis.ticks = element_blank())
  )


```

### Maps VIF

```{r column_inventory_maps_vif}

cant_inv_all_long_sub %>%
  filter(rarefication_threshold == 10) %>% 
  group_split(eras) %>%
  # head(1) %>%
  map(
    ~ p_map_cant_inv_offset(
      df = .x,
      var = "delta_cant_offset",
      # col = "divergent",
      subtitle_text = paste("VIF test | Eras:", unique(.x$eras)),
      breaks = seq(-8,8,1)
    ) +
      facet_grid(vif_max ~ estimate) +
      theme(axis.text = element_blank(),
            axis.ticks = element_blank())
  )


```


### Bias correlation

```{r inv_bias_plots_rarefication, fig.asp=1}

axis_limit <- 10

cant_inv_all_sub %>%
  filter(vif_max == 1000) %>% 
  group_split(eras) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x, aes(emlr_bias, emlr_anom)) +
      geom_hline(yintercept = 0) +
      geom_vline(xintercept = 0) +
      geom_bin2d(binwidth = 1) +
      scale_fill_viridis_c(trans = "log10") +
      coord_equal(
        xlim = c(-axis_limit, axis_limit),
        ylim = c(-axis_limit, axis_limit)
      ) +
      labs(title = paste("Eras:", unique(.x$eras))) +
      facet_grid(rarefication_threshold ~ basin_AIP)
  )


cant_inv_all_sub %>%
  filter(vif_max == 1000) %>% 
  group_split(eras) %>%
  # head(1) %>% 
  map(
    ~ ggplot(data = .x, aes(emlr_anom, nss_mod_truth)) +
      geom_hline(yintercept = 0) +
      geom_vline(xintercept = 0) +
      geom_bin2d(binwidth = 1) +
      scale_fill_viridis_c(trans = "log10") +
      coord_equal(
        xlim = c(-axis_limit, axis_limit),
        ylim = c(-axis_limit, axis_limit)
      ) +
      labs(title = paste("Eras:", unique(.x$eras))) +
      facet_grid(rarefication_threshold ~ basin_AIP)
  )

cant_inv_all_sub %>%
  filter(vif_max == 1000) %>% 
  group_split(eras) %>%
  # head(1) %>% 
  map(
    ~ ggplot(data = .x, aes(nss_mod_truth, nss_projection)) +
      geom_hline(yintercept = 0) +
      geom_vline(xintercept = 0) +
      geom_bin2d(binwidth = 1) +
      scale_fill_viridis_c(trans = "log10") +
      coord_equal(
        xlim = c(-axis_limit, axis_limit),
        ylim = c(-axis_limit, axis_limit)
      ) +
      labs(title = paste("Eras:", unique(.x$eras))) +
      facet_grid(rarefication_threshold ~ basin_AIP)
  )

```


## Zonal sections

### Sections

```{r cant_sections_rarefication_vif, fig.asp=0.5, eval=FALSE}

cant_section_all_long_sub %>%
  filter(vif_max == 1000) %>% 
  group_split(eras, rarefication_threshold, estimate) %>%
  head(1) %>%
  map(
    ~ p_section_zonal(
      df = .x,
      var = "delta_cant_offset",
      plot_slabs = "n",
      col = "divergent",
      subtitle_text = paste("Eras:", unique(.x$eras),
                            "| MLR Basins:", unique(.x$MLR_basins),
                            "| Estimate:", unique(.x$estimate)),
      breaks = c(-Inf,seq(-8,8,1), Inf)
    )
  )


```


```{r cant_sections_rarefication_facet}

legend_title <- expression(atop(Delta * C[ant, pos],
                               (mu * mol ~ kg ^ {
                                 -1
                               })))

cant_section_all_long_sub %>%
  filter(vif_max == 1000) %>% 
  group_split(eras, basin_AIP) %>% 
  # head(1) %>%
  map(
    ~ ggplot(data = .x) +
      geom_contour_filled(aes(lat, depth, z = delta_cant_offset),
                          breaks =c(-Inf,seq(-8,8,1), Inf)) +
      scale_fill_scico_d(
        palette = "vik",
        drop = FALSE,
        name = legend_title
      ) +
      guides(fill = guide_colorsteps(barheight = unit(8, "cm"),
                                     show.limits = TRUE)) +
      # scale_y_sqrt() +
      scale_y_continuous(trans = trans_reverser("sqrt"),
                         breaks = c(100,500,seq(1000,5000,1000))) +
      #scale_y_reverse() +
      scale_x_continuous(breaks = seq(-100, 100, 40),
                         limits = c(-85, 85)) +
      facet_grid(rarefication_threshold ~ estimate) +
      coord_cartesian(expand = 0) +
      labs(title = paste("Rarefication test | Eras", unique(.x$eras),
                         "| Basin:", unique(.x$basin_AIP)))
  )


```

```{r cant_sections_vif_facet}

cant_section_all_long_sub %>%
  filter(rarefication_threshold == 10) %>% 
  group_split(eras, basin_AIP) %>% 
  # head(1) %>%
  map(
    ~ ggplot(data = .x) +
      geom_contour_filled(aes(lat, depth, z = delta_cant_offset),
                          breaks =c(-Inf,seq(-8,8,1), Inf)) +
      scale_fill_scico_d(
        palette = "vik",
        drop = FALSE,
        name = legend_title
      ) +
      guides(fill = guide_colorsteps(barheight = unit(8, "cm"),
                                     show.limits = TRUE)) +
      # scale_y_sqrt() +
      scale_y_continuous(trans = trans_reverser("sqrt"),
                         breaks = c(100,500,seq(1000,5000,1000))) +
      #scale_y_reverse() +
      scale_x_continuous(breaks = seq(-100, 100, 40),
                         limits = c(-85, 85)) +
      facet_grid(vif_max ~ estimate) +
      coord_cartesian(expand = 0) +
      labs(title = paste("VIF test | Eras", unique(.x$eras),
                         "| Basin:", unique(.x$basin_AIP)))
  )


```

### Bias correlation

```{r section_bias_plots_rarefication, fig.asp=1}

axis_limit <- 20

cant_section_all_sub %>%
  filter(vif_max == 1000) %>% 
  group_split(eras) %>%
  # head(1) %>%
  map(
    ~ ggplot(data = .x, aes(emlr_bias, emlr_anom)) +
      geom_hline(yintercept = 0) +
      geom_vline(xintercept = 0) +
      geom_bin2d(binwidth = 1) +
      scale_fill_viridis_c(trans = "log10") +
      coord_equal(
        xlim = c(-axis_limit, axis_limit),
        ylim = c(-axis_limit, axis_limit)
      ) +
      labs(title = paste("Eras:", unique(.x$eras))) +
      facet_grid(rarefication_threshold ~ basin_AIP)
  )


cant_section_all_sub %>%
  filter(vif_max == 1000) %>% 
  group_split(eras) %>%
  # head(1) %>% 
  map(
    ~ ggplot(data = .x, aes(emlr_anom, nss_mod_truth)) +
      geom_hline(yintercept = 0) +
      geom_vline(xintercept = 0) +
      geom_bin2d(binwidth = 1) +
      scale_fill_viridis_c(trans = "log10") +
      coord_equal(
        xlim = c(-axis_limit, axis_limit),
        ylim = c(-axis_limit, axis_limit)
      ) +
      labs(title = paste("Eras:", unique(.x$eras))) +
      facet_grid(rarefication_threshold ~ basin_AIP)
  )

cant_section_all_sub %>%
  filter(vif_max == 1000) %>% 
  group_split(eras) %>%
  # head(1) %>% 
  map(
    ~ ggplot(data = .x, aes(nss_mod_truth, nss_projection)) +
      geom_hline(yintercept = 0) +
      geom_vline(xintercept = 0) +
      geom_bin2d(binwidth = 1) +
      scale_fill_viridis_c(trans = "log10") +
      coord_equal(
        xlim = c(-axis_limit, axis_limit),
        ylim = c(-axis_limit, axis_limit)
      ) +
      labs(title = paste("Eras:", unique(.x$eras))) +
      facet_grid(rarefication_threshold ~ basin_AIP)
  )

```



## Budgets

### Regional

```{r budgets_regional_rarefication, fig.asp=1}

cant_budget_all_sub %>%
  filter(vif_max == 1000,
         inv_depth == params_global$inventory_depth_standard) %>% 
  ggplot(aes(as.factor(rarefication_threshold), cant_pos_total, fill = basin_AIP)) +
  scale_fill_brewer(palette = "PuBuGn") +
  geom_col(col = "black") +
  facet_grid(estimate ~ eras) +
  theme(axis.text.x = element_text(angle = 90))

```


```{r budgets_regional_vif, fig.asp=1}

cant_budget_all_sub %>%
  filter(rarefication_threshold == 10,
         inv_depth == params_global$inventory_depth_standard) %>% 
  ggplot(aes(as.factor(vif_max), cant_pos_total, fill = basin_AIP)) +
  scale_fill_brewer(palette = "PuBuGn") +
  geom_col(col = "black") +
  facet_grid(estimate ~ eras) +
  theme(axis.text.x = element_text(angle = 90))

```

